---
title: "NATURAL LANGUAGE PROCESSING"
author: "kipngenokoech"
date: "7/11/2021"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning = FALSE)
```


## introduction
First before engaging in any machine learning task is to load the required libraries and files into the working environment. 
```{r}
#loading libraries and files
library(tm)
library(SnowballC)
library(wordcloud)
library(RColorBrewer)
library(pdftools)
library(topicmodels)
library(qdap)

text_2014 <- pdf_text("2014.pdf")#reading files into our environment
text_2016 <- pdf_text("2016.pdf")
text_2018 <- pdf_text("2018.pdf")
text_2019 <- pdf_text("2019.pdf")
text_2020 <- pdf_text("2020.pdf")

#Exporting our dataframe into a file
write.table(text_2014,"C:/Users/USER/Desktop/pandas/Natural-Language-P-rocessing/enockact/text_2014.text")
write.table(text_2016,"C:/Users/USER/Desktop/pandas/Natural-Language-P-rocessing/enockact/text_2016.text")
write.table(text_2018,"C:/Users/USER/Desktop/pandas/Natural-Language-P-rocessing/enockact/text_2018.text")
write.table(text_2019,"C:/Users/USER/Desktop/pandas/Natural-Language-P-rocessing/enockact/text_2019.text")
write.table(text_2020,"C:/Users/USER/Desktop/pandas/Natural-Language-P-rocessing/enockact/text_2020.text")
```

Creating corpus from our files. Corpus is a collection of text documents which can are machine readable.
```{r}
#creating a corpus
name<-file.path("C:/Users/USER/Desktop/pandas/Natural-Language-P-rocessing/enockact")
length(dir(name))
dir(name)
docs<-Corpus(DirSource(name))
docs
```

```{r}
#text transformation
ocs<-tm_map(docs,tolower)# tranforms text to lower case
docs<-tm_map(docs,removeNumbers)#remove numbers within our corpus
docs<-tm_map(docs,removePunctuation)# remove all forms of punctuations
docs<-tm_map(docs,removeWords,stopwords("english")) #remove commons words like is,in and where
docs<-tm_map(docs,stripWhitespace)#remove the spaces left after eliminating punctuations 
docs<-tm_map(docs,stemDocument)
```


```{r}
#creating sparse matrix
dtm<-DocumentTermMatrix(docs)
dim(dtm)
#specify sparsity matrix
dtm<-removeSparseTerms(dtm,0.51)
```

```{r}
#naming rows
rownames(dtm)<-c("2014","2016" ,"2018","2019","2020")

inspect(dtm[1:5,1:6])
```

## modelling and evaluation
```{r}
freq<-colSums(as.matrix(dtm))
#examing the head and the tail
ord<-order(-freq)
freq[head(ord)]
freq[tail(ord)]
```
President Kenyatta frequently use the word `kenya` and  `worker`,`wit`,`student`,`parent` being the least words used.

```{r}
#identifying association of words using correlation
findAssocs(dtm,"economi",corlimit = 0.9)
```
 `Better`,`fellow`,`infrastructure` have a correlation of `1` with the word `economi` ass shown above.

```{r}
#visual portrayal

wordcloud(names(freq),freq,min.freq = 20,scale = c(3,.5),colors = brewer.pal(5,"Dark2"))
```
Word cloud is used to visualize the word frequencies.The size of each word depends on the frequency of occurrence in our document. `Kenyan` has the highest frequency.

```{r}
#barpot
freq<-sort(colSums(as.matrix(dtm)),decreasing = TRUE)
word_freq<-data.frame(word=names(freq),freq=freq)
word_freq<-word_freq[1:10,]
coul <- brewer.pal(5, "Set2") 
barplot(word_freq$freq,names=word_freq$word,col=coul,main = "Word frequency",xlab = "words",ylab = "counts",ylim = c(0,160))
```
The frequency of each word has been shown clearly by the bar graph ,`kenyan` having the highest frequency.

```{r}

#comapring 2014 and 2020 speeches
speech14<-data.frame(speech=text_2014)
sent14<-sentSplit(speech14,"speech")
sent14$year<-"2014"
speech20<-data.frame(speech=text_2020)
sent20<-sentSplit(speech20,"speech")
sent20$year<-"2020"
#concatenate the datasets
sentences<-rbind(sent14,sent20)

```

```{r}
#comparing polarity
pol<-polarity(sentences$speech,sentences$year)
pol
```
 2020 speech appears to be the longest with `329` sentences and `3668` words.
 
 
```{r}
plot(pol)
```
2014  speech had strong negative sentiments with standard polarity of `0.323`compared to 2020 with `0.232` standard mean polarity as shown  by the polarity plot.

```{r}
#Dispersion plot of the following selected words(economi,fellow,billion,busi)

dispersion_plot(sentences$speech,grouping.var=sentences$year,c("economi","fellow","billion","busi"),color="black",bg.color="white")
```
This is quite interesting as these topics were discussed early on in the 2014 speech 
but at the end in the 2020 speech.
